---
title: "Globaltest"
author: "John Zhuang"
date: "2024-06-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
library(globaltest)
library(ggplot2)
library(dplyr)
```

# 0. Data Preparation

```{r}
data.horizontal = readRDS("data/derived/data_horizontal_joined.rds")
```

Keep only columns relevant for globaltest (all numerical datasets) and convert them into matrices:

```{r}
gt.matrix = as.matrix(data.horizontal[,1:701])
```

# 1. Globaltest

```{r}
Y = data.horizontal$vital_status
```
```{r, fig.width=70, fig.height=10}
x.mRSG = gt.matrix[, 1:470]
gt.mRSG = covariates(gt(Y ~ x.mRSG, model = "logistic"),
                     cluster = T, help.lines = T, alias = colnames(x.mRSG),
                     cex.labels = 0.5, colors = c(2,3))
```
```{r, fig.width=20, fig.height=10}
x.R2Gn = gt.matrix[, 471:668]
gt.R2Gn = covariates(gt(Y ~ x.R2Gn, model = "logistic"),
                     cluster = T, help.lines = T, alias = colnames(x.R2Gn),
                     cex.labels = 0.5, colors = c(2,3))
```
```{r, fig.width=15, fig.height=10}
x.RPPA = gt.matrix[, 669:701]
gt.RPPA = covariates(gt(Y ~ x.RPPA, model = "logistic"),
                     cluster = T, help.lines = T, alias = colnames(x.RPPA),
                     cex.labels = 0.7, colors = c(2,3))
```

## A. Zoomed-in miRNASeqGene and RNASeq2GeneNorm

Since the plots for miRNASeqGene and RNASeq2GeneNorm contain too many variables,
we will zoom in on the ones with the smallest p values.

```{r}
# Return the top most significant results from a Globaltest covariate object
# Select all significant p values with a limit of 0.05 (default), or any other limit
# or select the top n most significant variables (overwrites limit if not NA)
get.significant.pvalues = function(gtcov, limit = 0.05, n = NA) {
  res = result(gtcov)[,c("alias", "p-value")]
  dir = ((extract(gtcov))@extra)[,c("alias", "direction")]
  out = right_join(res, dir, by = "alias")
  names(out) = c("var", "p.value", "direction")
  out = out[order(out$p.value),] # Order by p-value
  #out$p.1.minus = 1 - out$p.value # Use 1-p for plotting
  if (!is.na(n)) {
    out = out[1:n,]  # Select the first (value) variables with the smallest p-values
  } else {
    out = out[out$p.value <= limit,] # Cut off based on p-value limit
  }
  out
}
```

```{r}
# Manually visualize the dataframe selected from get.significant.pvalues
# Optional: horizontal line drawing at selected p value.
gtcov.manual = function(gtp, limit = NULL) {
  p = ggplot(data = gtp, aes(x=var, y=log10(p.value), fill = direction)) +
    geom_bar(stat="identity", show.legend = T) + 
    scale_y_reverse(labels = c(0.001, 0.01, 0.1, 1)) +
    scale_x_discrete(guide = guide_axis(angle = 90)) +
    #theme(axis.text.x=element_text(size=rel(0.5), angle=90)) +
    scale_fill_manual(values = c(3,2)) +
    #scale_y_log10()
    ylab("p-value")
  
  if (!is.null(limit)) {
    p = p + geom_hline(yintercept=log10(limit), linetype="dashed", 
                color = "black")
  }
  p
}
```

For miRNASeqGene, we plot all variables that correspond to a significant p value (<0.05).

For RNASeq2GeneNorm, we plot the top 70 smallest p values (not all of them are significant).

The threashold alpha = 0.05 is plotted as a dashed black line.

```{r, fig.width=15, fig.height=10}
gtcov.manual(get.significant.pvalues(gt.mRSG), limit = 0.05)
gtcov.manual(get.significant.pvalues(gt.R2Gn, n = 70), limit = 0.05)
```

## B. Save significant (P <= 0.05) to external file

```{r}
results.globaltest = list("miRNASeqGene" = get.significant.pvalues(gt.mRSG),
                          "RNASeq2GeneNorm" = get.significant.pvalues(gt.R2Gn),
                          "RPPAArray" = get.significant.pvalues(gt.RPPA))
saveRDS(results.globaltest, file = "data/sel_features/results_globaltest_sig.rds")
```

## C. Save Globaltest models to external file

```{r}
models.globaltest = list("miRNASeqGene" = gt.mRSG,
                          "RNASeq2GeneNorm" = gt.R2Gn,
                          "RPPAArray" = gt.RPPA)
saveRDS(models.globaltest, file = "data/out_models/results_globaltest.rds")
```

